---
title: "3_FlaB-like flagellins in human gut metagenome"
author: "Andrea Borbon"
output: github_document
---

This notebook describes how we obtained the intersect between flagellins present in human gut metagenomes and FlaB-like candidates previously selected

#### Load libraries
```{r}
library(tidyverse)
library(tidytext)
library(conflicted)
```

```{r setup}
knitr::opts_knit$set(root.dir = "/ebio/abt3_projects/small_projects/aborbon/TLR5/SilentFlagellins_code/")
```

```{r}
conflict_prefer("filter","dplyr")
```

#### 1. Import files
##### 1.1 Import the list of accession IDs of flab-like flagellins with non-R/K in residue 467
```{r}
#This list was produced with Geneious Pro after manually checking the flagellin alignments and selecting those accessions that did not have R/K in the position 467
tofilter=read.delim("data/4_ToFilterFlaBFliCfinalhits.txt",header=F)
names(tofilter)="Accession"
tofilter

```

##### 1.2. Import Franzosa mapping results
```{r}
flagellin_diamond_franzosa=read.delim("data/mergedFlagFranzosa.txt",header=F,sep="")
names(flagellin_diamond_franzosa)=c("Study","Sample","qseqid","sseqid","pident","length","mismatch","gapopen","qstart","qend","sstart","send","evalue","bitscore")

```

##### 1.3. Import Lloyd-Price mapping results
```{r}
#Import diamond blastx results - reads mapped to flagellin accessions  
flagellin_diamond_lloyd=read.delim("data/mergedFlagellinLloyd.txt",header=F,sep="")%>%
  mutate(Study="LloydPrice_2019")
names(flagellin_diamond_lloyd)=c("Sample","qseqid","sseqid","pident","length","mismatch","gapopen","qstart","qend","sstart","send","evalue","bitscore","Study")
```

##### 1.4. Merge LloydPrice and Franzosa results
```{r}
mg.mapping.results=rbind(flagellin_diamond_franzosa,flagellin_diamond_lloyd)%>%
  rename(Accession=sseqid)

mg.mapping.results=mg.mapping.results%>%
  group_by(Accession)%>%
  count()

mg.mapping.results %>%
  head()
```

#### 2. Intersect of FlaB-like in human metagenomes
Here all flagellins present in human metagenomes will be intersected with the set of the FlaB-like flagellins from 1.1
```{r}
intersect.flab.mgs=inner_join(tofilter,mg.mapping.results)
nrow(intersect.flab.mgs)
  
intersect.flab.mgs %>%
  head()
```

#### 3. Export accessions of FlaB-like flagellins present in human metagenomes
This list will we used to pull out FlaB-like flagellins present in human metagenomes from the comlpete flagellin database
```{r}
write_lines(intersect.flab.mgs$Accession,"data/intersect.flab.metagones_accn.tsv")
```

#### 4. Remove FlgL sequences and de-replicate final sequences
```{r}
#subset intersect.flab.lloyd from flagellinDB
# seqtk subseq /ebio/abt3_projects/small_projects/aborbon/TLR5/Dalong_Flagellin/flic_flab_search/Dalong_flagellinDB_100 /ebio/abt3_projects/small_projects/aborbon/TLR5/Dalong_Flagellin/flic_flab_search/diamond/1_flab_2_flic/intersect.flab.metagones_accn.tsv > intersect.flab.metagones_accn.fasta
# 
# #Manually removed all proteins annotated as "FlgL"
# grep "FlgL" intersect.flab.metagones_accn.fasta | cat > toRemove.tsv
# seqtk subseq intersect.flab.metagones_accn.fasta toRemove.tsv > finalhits_ibd_tlr5epitope_flab_flic.fasta #Revisar aquí como uso seqtk para esto
# #This reduced the dataset to 129 sequences
```

#### 5. Dereplication of final sequences
Dereplication using CD-HIT at 0.99 identity.
```{r}
#Dereplication CD-HIT
# cd-hit -i finalhits_ibd_tlr5epitope_flab_flic.fasta -o derep0.99.flablike.humanmetagenomes.fasta -c 0.99 -T 10 -M 16000 -n 5
```


```{r}
sessionInfo()
```

